---
title: "ApoAnalysis"
author: "Hayley Brackenridge"
date: '2019-03-22'
output: html_document
---

## Data manipulation
```{r Data manipulation}
library(plyr)
apo<- read.csv("Mod_F_Harvestv2.csv")
str(apo)
Apo<- dplyr::select(apo, Population, Family, Individual, Bag, Tot.Seed, Num.Inflo.w.Seed)
ApoB<- subset(Apo, Bag=="Y") # subset for bagged individuals
# remove pop'ns w/ small num
ApoB <- ApoB[ !(ApoB$Population %in% "MA-GA"), ]
ApoB <- ApoB[ !(ApoB$Population %in% "MR-NC"), ]
ApoB <- ApoB[ !(ApoB$Population %in% "WN-NC"), ]

# add lat
lat<- read.csv("Population Locations.csv") # read in lat datasheet
lat<- dplyr::select(lat, Population, Latitude)
ApoBlat <- join(ApoB, lat, by = c("Population")) # combine lat & apo datasheets
str(ApoBlat)

# make a col saying Y/N seed production
for (i in 1:length(ApoBlat$Population)) {
  if (paste(ApoBlat$Tot.Seed[i])>0) {
    ApoBlat$SeedYN[i]<- paste("Y")
  }
  else
    ApoBlat$SeedYN[i]<- paste("N")
}

ApoBlat <- transform(ApoBlat, Population=reorder(Population, Latitude)) 
```

## Graph of proportion of seed production per pop
```{r Prop Seed per pop'n}
library(ggplot2)
# plot both proportion of seed producers & not seed producers in bargraph, organized by lat
Seed<- ggplot(ApoBlat,aes(x = Population,fill = SeedYN)) + 
  geom_bar(position = "fill") +
  labs(y="Seed Production")+
  scale_fill_manual(values = c("#C57B57","#041F1E")) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_text(size=12),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size=10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(fill=NA),
    plot.margin=unit(c(0.5,1,0,0), "cm")
  )
```

## Graph of av amount of seed produced per pop'n
```{r Seed per pop'n}
ApoBY<- subset(ApoBlat, SeedYN=="Y", na.rm=TRUE) #subset seed producers only
ApoBY$Num.Inflo.w.Seed # this produces NAs ?????
seedperinf<-ApoBY$Tot.Seed/ApoBY$Num.Inflo.w.Seed

# boxplot of seed count per pop'n
SperI<- qplot(x=Population, y=seedperinf, data=ApoBY, geom="boxplot", fill=I("#041F1E"), na.rm=T) + 
  labs(y="Number of seed per \n seed producing inflorescence") +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_text(size=12),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size=10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(fill=NA),
    plot.margin=unit(c(0.5,1,0,0), "cm")
  ) 
```

## seed production per unit length of exp inflo

```{r Seed per unit Inflo}
seedprod<- read.csv("data/F_Seed_Production.csv")
str(seedprod)
seedprodlat<- join(seedprod, lat, by = c("Population"))
seedprodlat<- transform(seedprodlat, Population=reorder(Population, Latitude)) 

### Not using this one
qplot(x=Population, y=(num.seed/inflo.length), data=seedprod, geom="boxplot") + 
  theme_classic()
histogram(~(num.seed/inflo.length), data=seedprod, type="count")
```

## criteria for apo

Rate the follow criteria as 1=yes, 0=no. Individuals with a total score of 3 are suspected to be apomictic

Criteria:

  1. More than 1 inflo produced seed
  2. Seed production > num inflos producing seed (i.e. at least one inflo produced more than 1 seed)
  3. No reason to suspect contamination (i.e. painted inflo, tie)-> only evaluated on individuals that meet first 2 criteria
  4. Consistently high production of seed (i.e. more than 3 seed for each inflo)

Using excel, num of individuals suspected of being apomictic = 4

```{r Prob of Apo}
library(knitr)
apocrit<- read.csv("Mod_F_Harvestv3.csv")
str(apocrit)

# make column of scores based on what they got for each criteria
## A = no prob of apo (i.e. no seed production), D = high prob of apo
for (i in 1:length(apocrit$Individual)) {
  if (paste(sum(apocrit$crit1[i],apocrit$crit2[i],apocrit$crit3[i],apocrit$crit4[i]))==0){
    apocrit$Score[i]<- paste("A") # no seed or on inflo producing seed
  } else {
      if (paste(sum(apocrit$crit1[i],apocrit$crit2[i],apocrit$crit3[i],apocrit$crit4[i]))==1){
    apocrit$Score[i]<- paste("B") # more than one inflo producing seed
      } else {
    if (paste(sum(apocrit$crit1[i],apocrit$crit2[i],apocrit$crit3[i],apocrit$crit4[i]))==2){
    apocrit$Score[i]<- paste("C") # more than one seed on more than one inflo
    } else {
    if (paste(sum(apocrit$crit1[i],apocrit$crit2[i],apocrit$crit3[i],apocrit$crit4[i]))==3){
    apocrit$Score[i]<- paste("D") # no contamination
    } else {
      if (paste(sum(apocrit$crit1[i],apocrit$crit2[i],apocrit$crit3[i],apocrit$crit4[i]))==4){
        apocrit$Score[i]<- paste("E") # high seed production
      }
    }
  }
  }
    }
}

apocrit<- subset(apocrit, Bag=="Y") # remove non-bagged individuals
# remove these pop'ns b/c small sample size 
apocrit <- apocrit[ !(apocrit$Population %in% "MA-GA"), ]
apocrit <- apocrit[ !(apocrit$Population %in% "MR-NC"), ]
apocrit <- apocrit[ !(apocrit$Population %in% "WN-NC"), ]
# remove pop'ns where seed was not counted
apocrit<- apocrit[!is.na(apocrit$Tot.Seed),]

# make a col saying Y/N seed production
for (i in 1:length(apocrit$Population)) {
  if (paste(apocrit$Tot.Seed[i])>0) {
    apocrit$SeedYN[i]<- paste("Y")
  }
  else
    apocrit$SeedYN[i]<- paste("N")
}

apocritY<- subset(apocrit, SeedYN=="Y")

ggplot(apocritY, aes(x=Population, fill=Score)) +
  geom_bar(position="fill") +
  theme_classic()
```

```{r Combining Graphs}
apograph1<- ggplot(ApoBlat,aes(x = Population, fill = SeedYN)) + 
  geom_bar(position = "fill") +
  labs(y="Seed production in isolation",fill="Seed\nProduction")+
  scale_fill_manual(values = c("#E0FBFC","#253237"),labels=c("No","Yes")) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size=11),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size=9),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(fill=NA),
    plot.margin=unit(c(0.5,0,0,0), "cm")
  )

apograph2<- qplot(x=Population, y=(num.seed/inflo.length), data=seedprod, geom="boxplot", fill=I("#253237"), na.rm=T) + 
  labs(y="Number of seed per\ncm inflorescence") +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_text(size=11),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size=9),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(fill=NA),
    plot.margin=unit(c(0.5,0,0,0), "cm")
  ) 

apograph3<- ggplot(apocritY, aes(x=Population, fill=Score)) +
  geom_bar(position="fill") +
  scale_fill_manual(values = c("#E0FBFC","#C2DFE3","#9DB4C0","#5C6B73","#253237"),labels=c("No Criteria","Criteria 1","Criteria 2","Criteria 3","Criteria 4"))+
  labs(y="Proportion of individuals", fill="Criteria Met") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size=11),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size=9),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(fill=NA),
    plot.margin=unit(c(0.5,0,0,0), "cm")
  )
  
library(sjPlot)
library(cowplot)
prow <- plot_grid(apograph1 + theme(legend.position="none"),
           apograph2 + theme(legend.position="none"),
           apograph3 + theme(legend.position="none"),
           align = 'v',
           labels = c("A", "B", "C"),
           hjust = -1,
           nrow = 3
           )

legend1<- get_legend(apograph1+theme(legend.justification = "top"))
legend3<- get_legend(apograph3+theme(legend.justification="bottom"))

p <- plot_grid(prow, legend1, rel_widths = c(3, .5))
p
q <- plot_grid(p, legend3, rel_widths = c(3,0))
q
```

```{r rough notes}
ApoBlat$Population

EF<- subset(ApoBlat, Population=="EF-IL")
nrow(EF) # 12 individuals bagged
EFY<- subset(EF, SeedYN=="Y")
nrow(EFY) # 8 individuals produced seed
max(EFY$Tot.Seed)

EFc<- subset(apocrit, Population=="EF-IL")
grepl("A",EFc$Score)
7/8# % of seed producing individuals that meet crit1
8/12 # % bagged individuals that produce seed

count(EFc$Score=="A") # 7 individuals meet criteria 1
count(EFc$Score=="B")
count(EFc$Score=="C")
count(EFc$Score=="D")
count(EFc$Score=="E")

mean(EFc$Tot.Seed)
sum(EFc$Tot.Seed)

WA<- subset(ApoBlat, Population=="WA-NC")
nrow(WA)
WAY<- subset(WA, SeedYN=="Y")
nrow(WAY)
max(WAY$Tot.Seed)
7/12 # % bagged individuals that produce seed

WAc<- subset(apocrit, Population=="WA-NC")

count(WAc$Score=="A")
count(WAc$Score=="B")
count(WAc$Score=="C")
count(WAc$Score=="D")
count(WAc$Score=="E")
5/7 # % of seed producing individuals that meet crit1

mean(WAc$Tot.Seed)
sum(WAc$Tot.Seed)

DO<- subset(ApoBlat, Population=="DO-GA")
nrow(DO)
grepl("Y",DO$SeedYN)

DOc<- subset(apocrit, Population=="DO-GA")
grepl("A",DOc$Score)
sum(DOc$Tot.Seed)

UC<- subset(ApoBlat, Population=="UC-NC")
nrow(UC)
grepl("Y",UC$SeedYN)
5/7

UCc<- subset(apocrit, Population=="UC-NC")
grepl("A",UCc$Score)
count(UCc$Score=="A")
sum(UCc$Tot.Seed)
4/5

fam<- subset(apocrit, Bag=="Y") %>%
  group_by(Population, Family)

ApoBY<- subset(ApoBlat, SeedYN=="Y")
sum(grepl("CH-IL",ApoBY$Population))
4/7
ApoBY<- ApoBY[!is.na(ApoBY$Tot.Seed),]
nrow(ApoBY) #68
Apo2<- subset(ApoBY, Tot.Seed<=3)
nrow(Apo2) #32
32/68
is.na(ApoBY$Tot.Seed)

# mean seed production
mean(ApoBY$Tot.Seed)
```
Num families, individuals, SeedY, Tot/SeedY%
CH-IL = 4,5,1,20
CL-NC = 3,4,1,25
CO-GA = 2,3,1,33
DO-GA = 4,4,2,50
EF-IL = 9,12,8,67
GL-IL = 5,7,3,43
HE-IL = 6,7,2,29
JO-NC = 7,10,4,40
LB-NC = 5,7,5,71
LI-NC = 5,5,2,40
MD-IL = 7,11,7,64
ON-NC = 4,9,3,33
PH-NC = 5,6,3,50
RA-NC = 6,6,4,67
TA-NC = 6,9,4,44
UC-NC = 6,7,5,71
WA-NC = 7,12,7,58
WL-NC = 6,8,5,63
WO-GA = 5,7,4,57

## Seed Prod x Lat
```{r Seed Prod x Lat}
# total seed production
Sprodmod<- lm(Tot.Seed~Latitude, data=ApoBlat)
summary(Sprodmod)

# seed production per unit length
Slenmod<- lm(num.seed/inflo.length~Latitude, data=seedprodlat)
summary(Slenmod)

# seed prod YN
blah<- ApoBlat
blah$SeedYN<- gsub("Y","1",blah$SeedYN)
blah$SeedYN<- gsub("N","0",blah$SeedYN)
Synmod<- lm(SeedYN~Latitude, data=blah)
summary(Synmod)
```

```{r}

```

